---
output:
  pdf_document: default
  html_document: default
---

# Intro to Data Science - HW 2

##### Copyright 2022, Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva

```{r}
# Enter your name here: Alireza Zarrinmehr
```

### Attribution statement: (choose only one and delete the rest)

```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```

### Reminders of things to practice from last week: 
Assignment arrow:  		<- <br>
The combine command: 	c( ) <br>
Descriptive statistics: 	mean( ) sum( ) max( ) <br>
Arithmetic operators:		+ - * /<br>
Boolean operators:		>    <    >=    <=   ==   !=  <br>

**This Week:** Explore the **quakes** dataset (which is included in R). 


<br>
<br>

**Step 0:** Prepare the data <br>
A. Copy the **quakes** dataset into a new dataframe (call it **myQuakes**), so that if you need to start over, you can do so easily (by copying **quakes** into **myQuakes** again). Summarize the variables in **myQuakes** using an R command. Also, explore the structure of the dataframe - using another R command. In a brief comment, describe what this dataset is about based on your analysis of the variables in it. 

```{r}
# Copying the data set.
myQuakes <- quakes
# summarizing the variables in the data set.
summary(myQuakes)
# exploring the structure of data frame.
str(myQuakes)
# this data frame illustrates the latitude and longitute of the point that earthquakes happened. It also contains the magnitude and depth values and tells you which station recorded the data 
```

**Step 1:** Explore the earthquake magnitude variable called **mag**  <br>

A.	What is the average magnitude? Use mean() or summary():

```{r}
#getting the average of the magnitudes 
mean(myQuakes$mag)
```

B.	What is the magnitude of the smallest earthquake? Use min() or summary() and save the result in a variable called **minQuake**:

```{r}
# storing the minimum magnitude of earthquakes
minQuake <- min(myQuakes$mag)
# printing the minimum magnitude of earthquakes
minQuake
```

C.	What is the magnitude of the largest earthquake? Use max() or summary() and save the result in a variable called **maxQuake**:

```{r}
# storing the maximum magnitude of earthquakes
maxQuake <- max(myQuakes$mag)
# printing the maximum magnitude of earthquakes
maxQuake
```

D.	Output the **fourth row** of the dataframe:

```{r}
#loading the library for slice
library(dplyr)
# the fourth row of the data frame
myQuakes[4,]
#another way of getting the fourth row
slice(myQuakes,c(4))
```

E. Create a new dataframe, with only the rows where the **magnitude is greater than 4**. How many rows are in that dataframe (report this number using code, do not count by looking at the output):

```{r}
#Creating a new dataframe, with only the rows where the magnitude is greater than 4
BigQuakes <- filter(myQuakes, mag>4)
#number of rows in the datafram
count(BigQuakes)
```

F.	Using the **arrange()** function, create a **sorted dataframe** based on magnitude and store it in **quakeSorted1**. Explain how this function works.

```{r}
quakeSorted1 <- arrange(myQuakes,mag)
quakeSorted1
# arrange() orders the rows of a data frame by the values of selected columns.
```

G. Now create a second sorted dataframe called **quakeSorted2** using the **order()** function. Explain in a comment how the function works.

```{r}
quakeSorted2 <- myQuakes[order(myQuakes$mag),]
quakeSorted2
#order returns a permutation which rearranges its first argument into ascending or descending order then we load those permutations in the data set rows
```

H. Print the top rows of each of the sorted dataframes using the **head()** function and explain in a comment what you see - did the two sorting functions produce a similar result?

```{r}
head(quakeSorted1)
# both produce the similar result. However, in the second one we can the the row number of the original data set which contains the values

```

```{r}
head(quakeSorted2)
# both produce the similar result. However, in the second one we can the the row number of the original data set which contains the values

```

I.	What are the latitude and longitude of the quake reported by the largest number of stations?  

```{r}
select(filter(myQuakes, stations==max(myQuakes$stations)), lat, long)
```

J.	What are the latitude and longitude of the quake with the smallest magnitude?  

```{r}
select(filter(myQuakes,stations==min(myQuakes$stations)), lat, long)
```

**Step 3:** Using conditional if statements <br>

A.	 Test if **maxQuake** is greater than 6 (output  yes  or  no ) <br>
**Hint:** Try modifying the following code in R: 

```{r}
if  (maxQuake > 6) "yes" else "no"
```

B.	 Following the same logic, test if **minQuake** is less than 3 (output  yes  or  no ):

```{r}
if  (minQuake < 3) "yes" else "no"
```

