---
output:
  pdf_document: default
  html_document: default
---

# Intro to Data Science HW 3

##### Copyright 2022, Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva

```{r}
# Enter your name here: Alireza Zarrinmehr
```

### Attribution statement: (choose only one and delete the rest)

```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```

### Reminders of things to practice from last week: 
Make a data frame:		data.frame( ) <br>
Row index of max/min:	which.max( )  which.min( )<br>
Sort value or order rows: arrange( )	sort( )   order( )<br>
Descriptive statistics: 	mean( ) sum( ) max( ) <br>
Conditional statement:	if (condition)  true stuff  else  false stuff <br>

### This Week: 
Often, when you get a dataset, it is not in the format you want. You can (and should) use code to refine the dataset to become more useful. As Chapter 6 of Introduction to Data Science mentions, this is called ** data munging. ** In this homework, you will read in a dataset from the web and work on it (in a data frame) to improve its usefulness.

## Part 1: Use read_csv( ) to read a CSV file from the web into a data frame:

A.	Use R code to read directly from a URL on the web. Store the dataset into a new dataframe, called **testDF**. <br>
The URL is:    <br>
"https://data-science-intro.s3.us-east-2.amazonaws.com/NYS_COVID_Testing.csv" <br>
**Hint:** use read_csv( ), not read.csv( ). This is from the **tidyverse package**. Check the help to compare them.

```{r}
#Load the library
library(tidyverse)
#Read from the csv file on web and store the data into testDF
testDF <- read_csv("https://data-science-intro.s3.us-east-2.amazonaws.com/NYS_COVID_Testing.csv")
testDF

```

B.	Use View( ), head( ), and tail( ) to examine the **testDF** dataframe. <br>
**Add a block comment that briefly describes what you see.**

```{r}
#Show the whole table of data
view(testDF)
#Just show the first 6 rows of data table
head(testDF)
#Just show the last 6 rows of data table
tail(testDF)
# This data is about the tests that were taken. There are information about thr date, age group, number of positive cases, number of tests, and age categories.
```

## Part 2: Create new data frames based on a condition:

A.	Use the table( ) command to summarize the contents of the **AgeCategory** variable in **testDF**. <br>
**Write a comment interpreting what you see   how many age categories are there in the dataset and what is the proportion of observations in each?**

```{r}
#Show values of the age category column and number of occurrence 
table(testDF$AgeCategory)
#we can see that there are 4 age categories. We have 2010 rows that belong to category of children, 1345 rows that belong to category of middle-aged_adults, 2686 rows that belong to category of  senior_citizens , and 1342 rows that belong to category of  young_adults.
```

B. Terms like "senior citizens" can function as *othering* language which demeans the people it seeks to describe. We can use the **str_replace_all()** function from tidyverse to find all instances of **senior_citizens** in the **AgeCategory** variable and replace them with **older_adults**. 

 <br>
In this case, we want to search for **senior_citizens** and replace it with **older_adults** in **testDF$AgeCategory** - how can you use this information to overwrite the **AgeCategory** in the function below:

```{r}
#Load the library
library(tidyverse)
#Replace the values "senior_citizens" with "older_adults" 
testDF$AgeCategory <- str_replace_all(testDF$AgeCategory, "senior_citizens", "older_adults")
#Check if the values have got replaced.
head(testDF)
```

C.	Create a dataframe (called **olderAdults**) that contains only the rows (observations) for which the value in the **AgeCategory** variable (column) is **older_adults**. Hint: Use subsetting.

```{r}
#Take a subset of the data that shows the information just for the older adults
olderAdults <- data.frame(subset(testDF, AgeCategory == "older_adults"))
#Check the dataframe
head(olderAdults)
```

D.	Use the dim() command on **olderAdults** to confirm that the data frame contains **2,686** observations and **5** columns/variables.

```{r}
#Check the number of columns and rows
dim(olderAdults)
```

E.	Use **subsetting** to create a new dataframe that contains only the observations for which the value in the **AgeCategory** variable is **young_adults**. The name of this new df should be **youngAdults**.

```{r}
#Take a subset of the data that shows the information just for the younger adults
youngAdults <- data.frame(subset(testDF, AgeCategory == "young_adults"))
#Check the dataframe
head(youngAdults)
```

F. Create one last data frame which only contains the observations for **children** in the **AgeCategory** variable of **testDF**. Call this new df **childrenDF**.

```{r}
#Take a subset of the data that shows the information just for children
childrenDF <- data.frame(subset(testDF, AgeCategory == "children"))
#Check the dataframe
head(childrenDF)
```

## Part 3: Analyze the numeric variables in the **testDF** dataframe.

A.	How many **numeric variables** does the dataframe have? You can figure that out by looking at the output of **str(testDF)**. 

```{r}
str(testDF)
#We have two columns of numeric variables
```

B.	What is the average number of total daily tests? Hint: Can you think of a mathematical function we've come across before to use on the **TotalTests** variable?

```{r}
#Get the mean of total test column
mean(testDF$TotalTests)
```

C.	How many tests were performed in the row with the highest number of total daily tests? What age category do they correspond to? 

```{r}
#loading the library
library(dplyr)
#show the row corresponding to the highest number of daily test
testDF[which.max(testDF$TotalTests),]
#Get the age category
testDF[which.max(testDF$TotalTests),5]
```

D. How many positive cases were registered in the row with the highest number of positive cases? What age category do they correspond to?

```{r}
#Loading the library
library(dplyr)
#show the row corresponding to the highest number of positive test
filter(testDF,PositiveCases==max(PositiveCases))
#Get the age category
testDF[which.max(testDF$PositiveCases),5]
```

E.	What is the total number of positive cases in **testDF**?

```{r}
#Total number of positive cases 
sum(testDF$PositiveCases)
```

F. Create a new variable in **testDF** which is the ratio of **PostiveCases** to **TotalTests**. Call this variable **PositivityRate** and explain in a comment what information it gives us.

```{r}
#create a function to calculate the ratio
Posratio<-function(PostiveCases,TotalTests){
  ratio<-PostiveCases/TotalTests
  return(ratio)
}
#Store that ration in a new column
testDF$PositivityRate<-Posratio(testDF$PositiveCases, testDF$TotalTests)
head(testDF)

#This variable gives us information about how bad the situation is regarding the positive test cases.
```

G. What is the average positivity rate in **testDF**? Hint: Use the **mean()** function on the new variable you created in F.

```{r}
#Get the average positivity rate
mean(testDF$PositivityRate)
```

## Part 4:  Create a function to automate the process from F-G:

A.	The following function should work most of the time. Make sure to run this code before trying to test it. That is how you make the new function known to R. **Add comments to each line explaining what it does:**

```{r}
#Creates a function called calculatePositivity that gets a dataset as input
calculatePositivity <- function(dataset) {
  #Calculates the positivity rate and store it in a new column
  dataset$PositivityRate <- dataset$PositiveCases/dataset$TotalTests
  #Calculates the average positivity rate
  avePositivity <- mean(dataset$PositivityRate)
  #returns the average positivity rate
  return(avePositivity)
}
```

B.	Run your new function on the **testDF** dataframe. Is the output of the function consistent with the output in Step G above? Explain.

```{r}
#Returns the average positivity rate of testDF
calculatePositivity(testDF)
```

C. Run the function on the **olderAdults** df you created earlier.

```{r}
#Returns the average positivity rate of olderAdults
calculatePositivity(olderAdults)
```

D.	Run the function on the **youngAdults** df.

```{r}
#Returns the average positivity rate of youngAdults
calculatePositivity(youngAdults)
```

E. Lastly, run the posivity function on the **childrenDF** dataframe.

```{r}
#Returns the average positivity rate of childrenDF
calculatePositivity(childrenDF)
```

F. In a comment, describe what you observe across these 3 datasets - which age group exhibits the highest positivity rate? How do these numbers compare to the **baseline** positivity rate in **testDF**?

```{r}
#Highest positivity rate belongs to young Adults at 0.05435627
#Lowest positivity rate belongs to children at 0.05034919
#childrenDF is less than mean(PositivityRate)
#Mean of 0.05363633 for testDF >  Mean of 0.05034919 for childrenDF
#Mean of 0.05363633 for testDF <  Mean of 0.05435627 for youngAdults
#Mean of 0.05363633 for testDF <  Mean of 0.05386202 for olderAdults
```

