---
output:
  pdf_document: default
  html_document: default
---

# Intro to Data Science - HW 10

##### Copyright 2022, Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva

```{r}
# Enter your name here: Alireza Zarrinmehr
```

### Attribution statement: (choose only one and delete the rest)

```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```

**Association mining** can be applied to many data problems beyond the well-known example of **finding relationships between different products in customer shopping data**. In this homework assignment, we will explore **real data** from the banking sector and look for **patterns associated with the likelihood of responding positively to a direct marketing campaign and signing up for a term deposit with the bank (stored in the variable  y )**. <br>
You can find out more about the variables in this dataset here: https://archive.ics.uci.edu/ml/datasets/bank+marketing

## Part 1: Explore Data Set

A.	Read the contents of the following URL to a dataframe called **bank** <br>
https://intro-datascience.s3.us-east-2.amazonaws.com/bank-full.csv

**Hint**: Even though this is a .csv file, chances are R won t be able to read it in correctly using the read_csv() function. If you take a closer look at the contents of the URL file, you may notice each field is separated by a **semicolon** (;) rather than a comma. 

In situations like this, consider using either **read.csv()** or **read.table()**, with two additional parameters. *sep=";"* defines how the data is seperated (the default is a comma), and *header=TRUE* defines that there is a header line in the dataset. 

```{r}
#Import required library.
library(tidyverse)
#Read the csv file and store it in a dataset called bank.
bank <- read.csv(header=TRUE,sep=";",'https://intro-datascience.s3.us-east-2.amazonaws.com/bank-full.csv')
#Review bank.
head(bank)
```

B. Make sure there are **41,188** rows and **21** columns in your **bank** df.

```{r}
#Get the number of rows and columns.
dim(bank)
```

C.	Next, we will focus on some key factor variables from the dataset, and convert a few numeric ones to factor variables. Execute the following command.  Write a comment describing how the conversion for each numeric variable works and what are the variables in the resulting dataframe.

```{r}
bank_new <- data.frame(
                     #Change the type of job category to factor.
                     job=as.factor(bank$job),
                     #Change the type of marital status to factor. 
                     marital=as.factor(bank$marital),
                     #Change the type of housing loan status to factor. 
                     housing_loan=as.factor(bank$housing),
                     #Create a new variable showing if the person is young. Change the type of variable to factor.
                     young=as.factor(bank$age<median(bank$age)),
                     #Create a new variable showing if the person contacted more than once. Change the type of variable to factor.
                     contacted_more_than_once=as.factor(bank$campaign>1),
                     #Create a new variable showing if the person previously contacted. Change the type of variable to factor.
                     contacted_before_this_campaign=as.factor(bank$previous<0),
                     #Create a new variable showing if the term deposit sign-ups were successful. Change the type of variable to factor.
                     success=as.factor(bank$y))
head(bank_new)
```

D.	Count the number of successful term deposit sign-ups, using the table( ) command on the **success** variable.

```{r}
#Count the successful term deposit sign-ups and failed term deposit sign-ups.
 table(bank_new$success)
```

E.	Express the results of problem C as percentages by sending the results of the table( ) command into the prop.table( ) command.

```{r}
#Calculate the probability of successful term deposit sign-ups.
prop.table(table(bank_new$success))
```

F.	Using the same techniques, show the percentages for the **marital** and **housing_loan** variables as well.

```{r}
#Calculate the probability of a person being married, divorced, or single.
prop.table(table(bank_new$marital))
#Calculate the probability of having housing loans.
prop.table(table(bank_new$housing_loan))

```

## Part 2: Coerce the data frame into transactions

A.	Install and library two packages: **arules** and **arulesViz**.

```{r}
#Import required library.
library(arules)
#Import required library.
library(arulesViz)
```

B.	Coerce the **bank_new** dataframe into a **sparse transactions matrix** called **bankX**.

```{r}
#Converting the bank_new dataframe into a sparse transactions matrix
#using as() function and specifying "transactions" in the function argument
bankX <- as(bank_new, 'transactions')
#check structure of the new transactions dataframe.
str(bankX)
```

C.	Use the itemFrequency( ) and itemFrequencyPlot( ) commands to explore the contents of **bankX**. What do you see?

```{r}
itemFrequency(bankX)
itemFrequencyPlot(bankX, topN=15)
#The itemFrequencyPlot is used create an item frequency bar plot for inspecting 
#the item frequency distribution for objects based on itemMatrix
#The argument topN specifies to only plot the topN items with the highest item
#frequency or lift 
```

D.	This is a fairly large dataset, so we will explore only the first 10 observations in the **bankX** transaction matrix: 

```{r}
#Show the first 10 observations or transactions in the matrix.
inspect(bankX[1:10]) 
```

E. Explain the difference between **bank_new** and **bankX** in a block comment:

```{r}
#bank_new is a dataframe with 41188 observations and 7 variables 
#while bankX is a sparse transactions matrix created from bank_new
#In bankX each observation shows all the variables in one transaction.
```

## Part 3: Use arules to discover patterns

**Support** is the proportion of times that a particular set of items occurs relative to the whole dataset. <br>
**Confidence** is proportion of times that the consequent occurs when the antecedent is present. <br>

A.	Use **apriori** to generate a set of rules with support over 0.005 and confidence over 0.3, and trying to predict who successfully signed up for a term deposit. <br>
**Hint:** You need to define the **right-hand side rule (rhs)**.

```{r}
rules1 <- apriori(bankX,
 parameter=list(supp=0.005, conf=0.3),
 control=list(verbose=F),
 appearance=list(default="lhs",rhs=("success=yes")))
#The apriori function is used to generate association rules
#A support of 0.5% and confidence of 30% is specified in the arguments
#The rhs is specified as "success=yes", denoting successfully 
#signed up for a term deposit
```

B.	Use inspect() to review of the **ruleset**. 

```{r}
inspect(rules1)
#The inspect function displays the rulesets generated by the apriori function
#and stored in rules1 variable
```

C.	Use the output of inspect( ) or inspectDT( ) and describe **any 2 rules** the algorithm found.  

```{r}
library(tidyverse)
#inspectDT(rules1)

#2 of the best rules or rules that hold the most significance can be 
#interpreted by comparing the confidence and lift value of the rules
#a higher confidence as well as higher lift value is desired

#The rule - job= student,marital=single,young=TRUE
#has a high confidence and lift. The rule denotes that the algorithm predicts
#that young,single students are likely to be successfully get signed up
#for a term deposit

#The rule - job=student,marital=single,young=TRUE,contacted_before_this_campaign=
#FALSE, denotes that single,young students that were never contacted about the
#campaign before also are highly likely to be successfully signed up for a term 
#deposit
```

D. Generate a partition tree from the dataframe (not the transactions)

```{r}
library(rpart)
cartTree <- rpart(success ~., data = bank_new, method="class")
```

View the model (as a tree), and then explain with a block comment if this tree is helpful

```{r}
library(rpart.plot)
prp(cartTree, faclen = 0, cex = 0.8, extra = 1)
#It is not a helpful tree
```
 




