---
output:
  pdf_document: default
  html_document: default
---

# Intro to Data Science - Lab 1
##### Copyright 2022, Jeffrey Stanton and Jeffrey Saltz   Please do not post online.

## Week 1 - Introduction to R

```{r}
# Alireza Zarrinmehr: 
```

### Please include nice comments. <br>
 
### Attribution statement: (choose only one and delete the rest)

```{r}
# 1. I did this lab assignment by myself, with help from the book and the professor.
# 2. I did this lab assignment with help from the book and the professor and these Internet sources:
# 3. I did this lab assignment with help from <Name of another student> but did not cut and paste any code.
```

Each student should run R-Studio on their computer (or via https://rstudio.cloud/).
Remember that R should always be installed before R-Studio (if you are not using
rstudio.cloud or Google Colab)

1. Add together all the numbers between 1 and 10 (inclusive). Take note of the result. Remember, every student should type and run the code on their machine.

```{r}
1+2+3+4+5+6+7+8+9+10
```

2. Now	create	a	vector	of	data	that	contains	the	numbers	between	1	and	10
(inclusive).	Here	is	a	line	of	code	to	do	that: <br>

`myNumbers <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)`

```{r}
myNumbers <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
myNumbers
```

3. Now add together all of the numbers that are in the vector **myNumbers**. There is a built-in function within R that can do this for you in one step: Take a guess as to the name of that function and run it on **myNumbers**. Check your result against the results of question 1.

```{r}
sum(myNumbers)
```

4. R can do a powerful operation called ** vector math ** in which a calculation runs on every element of a vector. Try vector math on **myNumbers** by adding 10 to each element of myNumbers, and storing the result in **myNewNumbers**. Print out **myNewNumbers**.

```{r}
myNewNumbers <- myNumbers + 10
myNewNumbers

```

5. Efficiently calculate a sum of the numbers between 11 and 20 (inclusive), using techniques from the problems above. **Hint:** use `c(11:20)`

```{r}
sum(c(11:20))
sum(11:20)
```

6. Calculate a sum of all of the numbers between 1 and 100 (inclusive), using techniques from the problems above.

```{r}
sum(1:100)
```


7. Make sure you have a variable **myNumbers**, that is a vector of 10 numbers (1,2,3,4,5,6,7,8,9,10)

```{r}
myNumbers <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
myNumbers
```

8. Type the following commands in the code cells below and run each one: 

<br>`mean(myNumbers)` <br>`median(myNumbers)` <br>`max(myNumbers)` <br>`min(myNumbers)` <br>`length(myNumbers)`

```{r}
mean(myNumbers)
```

```{r}
median(myNumbers)
```

```{r}
max(myNumbers)
```

```{r}
min(myNumbers)
```

```{r}
length(myNumbers)
```

9. Repeat the commands from above, this time adding a comment to each line of code in your file explaining what it does. The comment character is  # .

```{r}
#calculating the mean value of the set of numbers between 1 to 10 
mean(myNumbers)
```

```{r}
#calculating the median value of the set of numbers between 1 to 10 
median(myNumbers)
```

```{r}
#calculating the maximum value of the set of numbers between 1 to 10 
max(myNumbers)
```

```{r}
#calculating the minimum value of the set of numbers between 1 to 10 
min(myNumbers)
```

```{r}
#calculating the distance between set of numbers between 1 to 10 
length(myNumbers)
```

10. In a comment, explain the output of the following command: <br> `myNumbers > 5`

```{r}
# returns a true value for any number greater than 5 and false value for every number less than 5
myNumbers>5
```

11. Explain what in is **bigNum** after executing the following command: 

`bigNum <- myNumbers[myNumbers > 5]`

```{r}
# returns a tuple including all the numbers that are grater than 5 from myNumbers
bigNum <- myNumbers[myNumbers > 5]
bigNum
```

12. Whenever you need R to explain what a command does and how it works, use the **?** command or the **help()** command. Add and run these commands: <br>
`?mean` <br>`help("mean")`

```{r}
?mean
```

```{r}
help("mean")
```
