---
output:
  pdf_document: default
  html_document: default
---

# Intro to Data Science - Lab 7
##### Copyright 2022, Jeffrey Stanton and Jeffrey Saltz   Please do not post online.

## Week 7 - Using	ggplot	to	Build	 Complex	Data	Displays

```{r}
# Enter your name here: Alireza Zarrinmehr
```

### Please include nice comments. <br>
### Instructions: 
Run the necessary code on your own instance of R-Studio.

### Attribution statement: (choose only one and delete the rest)

```{r}
# 1. I did this lab assignment by myself, with help from the book and the professor.
```

 Geology rocks but geography is where it's at. . .  (famous dad joke). In
a global economy, geography has an important influence on everything from
manufacturing to marketing to transportation. As a result, most data scientists will
have to work with map data at some point in their careers. <br><br>

An add-on to the **ggplot2** package, called **ggmap**, provides powerful tools for plotting and shading maps.<br><Br>
Make sure to install the **maps**, **mapproj**, and **ggmap** packages before running the following:



```
library(ggplot2); library(maps); library(ggmap); library(mapproj)
us <- map_data("state")
us$state_name <- tolower(us$region)
map <- ggplot(us, aes(map_id= state_name))
map <- map + aes(x=long, y=lat, group=group) +
geom_polygon(fill = "white", color = "black")
map <- map + expand_limits(x=us$long, y=us$lat)
map <- map + coord_map() + ggtitle("USA Map")
map
```



1. Paste the code below and add a comment for each line, explaining what that line of code does.

```{r}
library(ggplot2); library(maps); library(ggmap); library(mapproj)
#Read the maps package and put it in us
us <- map_data("state")
#Creating a new column state name. converting the rigon and make it lowercase.
us$state_name <- tolower(us$region)
#Start to create the plot. Give the map ID, State name.
map <- ggplot(us, aes(map_id= state_name))
#Group all the states together
map <- map + aes(x=long, y=lat, group=group) +
#Give the fill and color
geom_polygon(fill = "white", color = "black")
#Take my map and make sure it is expanded.
map <- map + expand_limits(x=us$long, y=us$lat)
#flatten the map and put it in scale.
map <- map + coord_map() + ggtitle("USA Map")
map
```

2. The map you just created fills in the area of each state in white while outlining it with a thin black line. Use the **fill=** and **color=** commands inside the call to **geom_polygon( )** to reverse the color scheme.<br><br>
Now paste and run the following code:<br><br>


```
ny_counties <- map_data("county","new york")
ggplot(ny_counties) + aes(long,lat, group=group) + geom_polygon(fill
= "white", color = "black")
```


```{r}
ny_counties <- map_data("county","new york")
ggplot(ny_counties) + aes(long,lat, group=group) + geom_polygon(fill
= "white", color = "black")
```

3. Just as in step 2, the map you just created fills in the area of each county in black
while outlining it with a thin white lines. Use the **fill=** and **color=** commands
inside the call to **geom_polygon( )** to reverse the color scheme.

```{r}
ny_counties <- map_data("county","new york")
ggplot(ny_counties) + aes(long,lat, group=group) + geom_polygon(fill
= "black", color = "white")
```

4. Run `head(ny_counties)` to verify how the county outline data looks

```{r}
head(ny_counties)

```

5. Make a copy of your code from step 3 and add the following subcommand to
your ggplot( ) call (don t forget to put a plus sign after the **geom_polygon( )** statement to tell R that you are continuing to build the command):<br><br>
`coord_map(projection = "mercator")` <br><br>
In what way is the map different from the previous map. Be prepared to explain
what a Mercator projection is. 

```{r}
ny_counties <- map_data("county","new york")
ggplot(ny_counties) + aes(long,lat, group=group) + geom_polygon(fill
= "black", color = "white") + coord_map(projection = "mercator")

#The map is wider in the last question.This looks a little tighter
```

6. Grab a copy of the nyData.csv data set from:
https://intro-datascience.s3.us-east-2.amazonaws.com/nyData.csv
<br><br>Read that data set into R with **read_csv()**. This will require you have installed
and libraried the **tidyverse** package. The next step assumes that you have named the
resulting data frame ** nyData. **

```{r}
library(tidyverse)
nyData <- read_csv("https://intro-datascience.s3.us-east-2.amazonaws.com/nyData.csv")
nyData
```

7. Next, merge your **ny_counties** data from the first set of
questions with your new **nyData** data frame, with this code:<br><br>
`mergeNY <- merge(ny_counties,nyData,all.x=TRUE,by.x="subregion",by.y="county")`



```{r}
mergeNY <- merge(ny_counties,nyData,all.x=TRUE,by.x="subregion",by.y="county")
```

8. Run `head(mergeNY)` to verify how the merged data looks.

```{r}
head(mergeNY)
```

9. Now drive the fill color inside each county by adding the **fill** aesthetic inside of
your **geom_polygon( )** subcommand (fill based on **pop2000**).

```{r}
ggplot(mergeNY) + aes(long,lat, group=group) + geom_polygon(aes(fill
= pop2000),  color = "black") + coord_map(projection = "mercator")

```
