---
output:
  pdf_document: default
  html_document: default
---

# Intro to Data Science - Lab 2
##### Copyright 2022, Jeffrey Stanton and Jeffrey Saltz   Please do not post online.

## Week 2 - Sorting Data and Ordering a Data Frame

```{r}
# Enter your name here: Alireza Zarrinmehr
```

### Please include nice comments. <br>

### Attribution statement: (choose only one and delete the rest)

```{r}

# 3. I did this lab assignment with help from <Saiteja Namani> but did not cut and paste any code.
```

1. Make	a	copy	of	the	built-in	iris	data	set	like	this:<br>
`myIris	<-	iris`

```{r}
#data() # reviewing all the data sets delivered with R
myIris	<-	iris
```

2. Get	an	explanation	of	the	contents	of	the	data	set	with	the	**help()**	function:<br>
`help("iris")`

```{r}
help("iris")
```

3. Explore	**myIris**	via the	**str()**	and	**glimpse()**	functions	(note:	you	need	to	install	and
library	** tidyverse **	to	use	**glimpse()**).	Which	do	you	think	is	better?	Why? Explain in a comment.

```{r}
library(tidyverse)
glimpse(myIris)
str(myIris)
# glimpse is better because it gives me more information
```

4. Summarize	the	variables	in	your	copy	of	the	data	set,	like	this:<br>
`summary(myIris)`

```{r}
summary(myIris)
```

5. The	**summary()**	command	provided	the	mean	of	each	numeric	variable.	Choose
the	variable	with	the	highest	mean and	list	its	contents.	Any
variable	can	be	echoed	to	the	console	simply	by	typing	its	name.	Here s	an
example	that	echoes	the	variable	with	the	lowest	mean:<br>
`myIris$Petal.Width`

```{r}
myIris$Sepal.Length
```

6. Now	sort	that attribute by	calling	the	**sort()**	function	and	supplying	that	variable.
Remember	to	choose	the	variable	with	the	highest	mean.

```{r}
sort(myIris$Sepal.Length)
```

7. Now	repeat	the	previous	command,	but	this	time	use	the	**order()**	function,	again
using the	variable	with	the	highest	mean.

```{r}
order(myIris$Sepal.Length)
```

8. Write	a	comment	in	your	R	code	explaining	the	difference	between	**sort()**	and
**order()**.	Be	prepared	to	explain	this	difference	to	the	class.

```{r}
#Sort puts all the values in increasing order and Order tells you which rows/indexes do these values belong to 
```

9. Now	use	the	**order()**	command	to	reorder	the	whole	data	frame, and	store	the	new
dataframe	in	a	variable	called	** sortedDF **:

```{r}
#view(myIris)
sortedDF <- myIris[order(myIris$Sepal.Length),]
sortedDF
view(sortedDF)
```

10. Now	sort	the	dataframe	using	**arrange()**,	which	is	part	of	the	**tidyverse**	package.
This	time,	sort	based	the attribute	with	the	lowest mean.	Store	the	new
dataframe	in	a	variable	called	** sortedDF1 **

```{r}
library(tidyverse)
sortedDF1 <- myIris %>% arrange(Petal.Width)
sortedDF1
```

11. Finally,	use	**head()**	to	examine	your	reordered	data	frames	and	be	prepared	to
report	on	the	first	few	rows


```{r}
head(sortedDF)
head(sortedDF1)
```

12. What	does	the	following	line	of	code	do?<br>
`myIris [	,	1]`

```{r}
myIris [	,	1]
#gives me all the rows for the first column.
```

13. What	is	the	difference	(if	any)	between:<br>
`myIris [	,	"Sepal.Length"]` and <br> 
`myIris $Sepal.Length`

```{r}
myIris [	,	"Sepal.Length"]
myIris $Sepal.Length
#there is no difference
```

14. Write	the	R	code	that	outputs	the	** Sepal.Length **	attribute	values,	using	the **select()**	command.

```{r}
select(myIris, Sepal.Length)
```

15. Create	a	new	column	(called	** Ave.Length **)	in	**myIris**,	which,	for	each	row,	is	the average	of	**Sepal.Length**	and	**Petal.Length**.

```{r}
myIris$Ave.Length <- (myIris$Sepal.Length+ myIris$Petal.Length)/2
myIris
```

16. What	does	the	following	line	of	code	do:<br>
`which.min(myIris$Petal.Length)`

```{r}
which.min(myIris$Petal.Length)
#gives me the index/row that contain the min value
myIris[23, ]
```

17. Using	the	code	from	the	previous	step,	output	the	row	(iris	observation)	with	the
smallest	petal	length.

```{r}

myIris[which.min(myIris$Petal.Length), ]
```

18. Create	a	new	dataframe,	with	just	the	**Petal.Length**	and	**Petal.Width** attributes

```{r}
newDF <- myIris[,c('Petal.Length', 'Petal.Width')]
newDF
```

19. Create	a	new	dataframe,	using	the	**slice()**	function,	with	only	the	first	three	rows
in	the	**myIris**	dataframe.

```{r}
library(tidyverse)
smallDF <- myIris%>%slice(1:3)
smallDF
```

20. Create a	new	dataframe,	which	is	a	subset	of	**myIris**,	that	only	includes	rows
where **Petal.Length is	less	than	1.4**,	store	in	**shortPetalDF**

```{r}
shortpetalDF <- myIris[myIris$Petal.Length<1.4,]
shortpetalDF

```

21. How	many	rows	are	in	the	**shortPetalDF**?

```{r}
nrow(shortpetalDF)
```

22. The	homework	asks	you	to	create	a	conditional	statement	with	**if**	and	**else**.	A
conditional	statement	is	part	of	a	larger	group	of	specialized	commands	that
control	the	 flow 	of	a	program	 	what	command	gets	run	and	when.	You	can	get
help	on	if,	else,	and	other	control	words.	Add	and	run	these	commands:<br>
`help("if")` <br>
`help("Control")`<br>
Now	add	and	run	your	first	conditional	statement:<br>
`if	(sum(myNumbers)	>	40)	print("The	sum	is	greater	than	40.")`

```{r}
help("if")
```

```{r}
help("Control")
```

```{r}
myNumbers <- c(2,5,4,7,8,20,9,3,1)
if(sum(myNumbers)>40) print("The sum is greater that 40") else print("No")
```

